The CrypTool Script: \\ Cryptography, Mathematics and More \\ \vspace{30pt} {\large Background reading\\%*[-8pt] for the free e-learning program CrypTool\\ (10th edition -- distributed with CrypTool version 1.4.30)} \vspace{60pt}

### 2009-11-27 script/en/chapters/numbertheory.tex
Script[de+en]: Many small changes, e.g. Sage samples collected in files, use of iload; described how to deal with Sage notebook using interact; substituted href by url.

\clearpage
% ++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\vspace{4ex}
\begin{center}
\fbox{\parbox{15cm}{
    \emph{Hermann Hesse\footnotemark:}\\
    To let the possible happen, you again and again have to try the impossible.
}}
\end{center}
\addtocounter{footnote}{0}\footnotetext{%
  Hermann Hesse, German/Swiss writer and Nobel Prize winner, July 2, 1877 $-$ August 9, 1962.}
% ++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++

@@

An exhaustive overview about the factoring records
\index{Factorization!factoring records} of composed integers using different
methods can be found on the following web pages:
\vspace{-10pt}
\begin{itemize}
\item[]
     \href{http://www.crypto-world.com}
  {\texttt{http://www.crypto-world.com}} \\
     \href{http://www.tutorgig.com/ed/RSA\_number}
  {\texttt{http://www.tutorgig.com/ed/RSA\_number}}  ~~ The RSA Factoring Challenge \\
     \href{http://en.wikipedia.org/wiki/Integer_factorization_records}
  {\texttt{http://en.wikipedia.org/wiki/Integer\_factorization\_records}}
\end{itemize}
%\footnote{%
%This site was not quite up-to-date end of May 2005: RSA-200 was missing.}

@@

\vskip +10pt
The last records\footnote{%
The 'RSA numbers' are certain large semiprime numbers (i.e., numbers with
exactly two prime factors)\index{Number!semi prime}. 
They were generated and published by the 
company RSA Security and they form the basis of the RSA Factoring 
Challenge, in which factorizations for these numbers are sought.
See \href{http://www.rsa.com/rsalabs/node.asp?id=2092}
 {\texttt{http://www.rsa.com/rsalabs/node.asp?id=2092}}.

The first RSA Factoring Challenge labeled the numbers, from RSA-100 
to RSA-500, according to their number of decimal digits; the second 
RSA Factoring Challenge labeled the numbers according to their number 
of binary digits. Within the second challenge cash prizes have been 
offered for successful factorizations of RSA576 to RSA2048 (RSA576, 
RSA640 etc. using 64 bit steps upwards).
An exception to this is RSA617, which was created prior to the 
change in the numbering scheme.

@@

The 'C-numbers' originate from the Cunningham project:
\href{http://www.cerias.purdue.edu/homes/ssw/cun/}
     {\texttt{http://www.cerias.purdue.edu/homes/ssw/cun/}}
                                                               }
with factorization algorithms for composed numbers are
listed in the table~\ref{factorizationrecords}.

\begin{table}[ht]
\begin{center}
\begin{tabular}{|c|cccc|}
\hline
	& {\bf Decimal digits} & {\bf Binary digits} & {\bf Factored on} & {\bf Factored by} \\
\hline
	C307 & 		307 & 1017 & May, 2007 & Jens Franke et al. \\
	RSA-200 & 	200 & 663 & May, 2005 & Jens Franke et al. \\
	RSA640\footnotemark & 	193 & 600 & November, 2005 & Jens Franke et al. \\
	C176 & 		176 & 583 & May, 2005 & Kazumaro Aoki et al. \\
	RSA576 & 	174 & 576 & December, 2003 & Jens Franke et al. \\
	RSA-160 & 	160 & 530 & April, 2003 & Jens Franke et al. \\
	C158 & 		158 & 523 & January, 2002 & Jens Franke et al. \\
	RSA-155	&	155 & 512 & August, 1999 & Herman te Riele et al. \\
\hline
\end{tabular}
\caption{The current factoring records (as of July 2009)}    % Eyecatcher_neue-Mersenne 
\label{factorizationrecords}
\end{center}
\end{table} 
\footnotetext{%
A research group of the GISA solved this challenge which was awarded with
20,000 US dollar using the GNFS method. The researchers needed about five
months to divide this number into its both 320 bit long prime factors.\\
RSA Labs offers its challenges since the beginning of the 1990th.
The next bigger challenge awarded with 30,000 US dollars is RSA-704.\\
See:
\href{http://www.heise.de/newsticker/meldung/print/65957}
{\texttt{http://www.heise.de/newsticker/meldung/print/65957}}
}

@@

-----------Testcase: (1000, 10000) [Should deliver 3.260.044]
Number of private keys for modulus in a given range: 3260044 (comb=2312),
Number of primes in a given range: 1061
Time: CPU 0.63 s, Wall: 0.66 s
\end{Verbatim}
\caption{How many RSA keys are there knowing the range of the public key n?}
\label{nt_sagesample_Count_RSA_Keys}
\end{sagecode}

>>>>

\clearpage
% ++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\vspace{4ex}
\begin{center}
\fbox{\parbox{15cm}{
    \emph{Hermann Hesse\footnotemark:}\\
    To let the possible happen, you again and again have to try the impossible.
}}
\end{center}
\addtocounter{footnote}{0}\footnotetext{%
  Hermann Hesse, German/Swiss writer and Nobel Prize winner, July 2, 1877 $-$ August 9, 1962.\\}
% ++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++

@@

An exhaustive overview about the factoring records
\index{Factorization!factoring records} of composed integers using different
methods can be found on the following web pages:
\vspace{-10pt}
\begin{itemize}
\item[]
     \url{http://www.crypto-world.com} \\
     \url{http://www.tutorgig.com/ed/RSA\_number}  ~~ The RSA Factoring Challenge \\
     \url{http://en.wikipedia.org/wiki/Integer_factorization_records}
\end{itemize}
%\footnote{%
%This site was not quite up-to-date end of May 2005: RSA-200 was missing.}

@@

\vskip +10pt
The last records\footnote{%
The 'RSA numbers' are certain large semiprime numbers (i.e., numbers with
exactly two prime factors)\index{Number!semi prime}. 
They were generated and published by the 
company RSA Security and they form the basis of the RSA Factoring 
Challenge, in which factorizations for these numbers are sought.\\
See \url{http://www.rsa.com/rsalabs/node.asp?id=2092}.

RSA Labs offers its challenges since the beginning of the 1990th.
The first RSA Factoring Challenge labeled the numbers, from RSA-100 
to RSA-500, according to their number of decimal digits; the second 
RSA Factoring Challenge labeled the numbers according to their number 
of binary digits. Within the second challenge cash prizes were 
offered for successful factorizations of RSA576 to RSA2048 (RSA576, 
RSA640 etc. using 64 bit steps upwards --- An exception to this is
RSA617, which was created prior to the change in the numbering scheme).\\
But the RSA challenges ended ahead of time in 2007, RSA Inc. retracted
the prize.
  \index{Challenge}
  \index{Cipher challenge}
  \index{Crypto challenge}

The 'C numbers' originate from the Cunningham project:
\url{http://www.cerias.purdue.edu/homes/ssw/cun/}.
                          }
with factorization algorithms for composed numbers are
listed in the table~\ref{factorizationrecords}.

\begin{table}[ht]
\begin{center}
\begin{tabular}{|c|cccc|}
\hline
	& {\bf Decimal digits} & {\bf Binary digits} & {\bf Factored on} & {\bf Factored by} \\
\hline
	C307 & 		307 & 1017 & May, 2007 & Jens Franke et al. \\
	RSA-200 & 	200 & 663 & May, 2005 & Jens Franke et al. \\
	RSA640\footnotemark & 	193 & 600 & November, 2005 & Jens Franke et al. \\
	C176 & 		176 & 583 & May, 2005 & Kazumaro Aoki et al. \\
	RSA576 & 	174 & 576 & December, 2003 & Jens Franke et al. \\
	RSA-160 & 	160 & 530 & April, 2003 & Jens Franke et al. \\
	C158 & 		158 & 523 & January, 2002 & Jens Franke et al. \\
	RSA-155	&	155 & 512 & August, 1999 & Herman te Riele et al. \\
\hline
\end{tabular}
\caption{The current factoring records (as of July 2009)}    % Eyecatcher_neue-Mersenne 
\label{factorizationrecords}
\end{center}
\end{table} 
\footnotetext{%
A research group of the GISA solved this challenge which was awarded with
20,000 US dollar using the GNFS method. The researchers needed about five
months to divide this number into its both 320 bit long prime factors.

@@

The next bigger challenge is RSA-704.
See \url{http://www.heise.de/newsticker/meldung/print/65957}.
}

@@

-----------Testcase: (1000, 10000) [Should deliver 3.260.044]
Number of private keys for modulus in a given range: 3260044 (comb=2312),
Number of primes in a given range: 1061
Time: CPU 0.63 s, Wall: 0.66 s
\end{Verbatim}
\caption{How many RSA keys are there if you know a range for the public keys n?}
\label{nt_sagesample_Count_RSA_Keys}
\end{sagecode}
